Future challenges for Research Software Engineering

### 2025-05-19 future_challenges_for_RSEng.tex
Language edits

So where do RSEs fit into this picture? Even today we expect to see RSEs providing technical skills and input
to research activities in a range of different areas, these include:
high-quality, robust architecture and design for research software;
seamless integration with research and data infrastructures;
development of software for novel and emerging architectures;
application of specialist numerical and statistical methods within software;
optimisation of codes for use on high-performance computing infrastructure;
and green computing - efficient implementations of algorithms/methods and use of hardware

@@

\subsection{Definition and Evolution of RSEs}
\begin{whatis}{}{definition-evolution-rse}
The term RSE is established in the minds of university boards and leadership, as a valuable member in the academic framework.
New RSEs are trained through a specialised Master's programme, and their support is valued by researchers. We still teach researchers basic software skills, but the content of this lecture has changed drastically.
\end{whatis}
% Here we consider how the topics of RSEng will change in the face of the digitalisation of society.
% Also here: What will our topics be after version control doesn't need to be taught anymore?
% Digitalisation will also feature more heavily in the domain curricula.
In this discussions the workshop participants found that
the digital skills of domain scientists will increase,
and software will become easier to interact with,
reducing reliance on RSEs to train newcomers on the most basic tasks.
This means, that today's workshop content of proper versioning and data management could become obsolete
by using more intuitive tools.
Introductory courses can then focus on more higher-level concepts.
User training sessions might focus on more advanced topics
that cater to a smaller audience, or involve mentoring more frequently.

@@

\subsection{Development of the RSE community}
\begin{whatis}{}{development-rse-community}
National RSE communities are increasingly collaborating with local partners, fostering stronger networks across countries. Regular international conferences further amplify these connections, promoting the global exchange of ideas. The introduction of RSE Master's programmes, coupled with a shift in culture towards valuing Research Software, has significantly elevated its profile. This growing recognition has led to a surge in membership of national societies, reflecting the increasing importance of RSEng in the research landscape.
\end{whatis}
In recent years, we have seen the formation of RSE institutes and advocacy groups, such as
the UK Software Sustainability Institute\footnote{\url{https://www.software.ac.uk}},
the US Research Software Sustainability Institute\footnote{\url{https://urssi.us}},
the European Virtual Institute for Research Software Excellence\footnote{\url{https://everse.software}},
and the EOSC Opportunity Area Expert Group 7 on Research Software%
\footnote{\url{https://eosc.eu/opportunity-area-exp/oa7-research-software}}.
We expect these initiatives to help increase the visibility and accelerate
the professionalisation%
of the RSE role.
In addition, a multitude of RSE communities self-organised at the national
and regional level to support more viable RSE career paths in academia.
We can expect RSE positions to become more easily accessible, possibly through centrally-funded long-term positions at research institutions.
The community of RSE's will get a significant boost in their acceptance if RSE Master's programmes
become available at various universities.
With increased numbers, the visibility of the RSE community to industry will be enhanced.
\begin{story}{Development of RSE community}{RSEcommunity}
Kay stands proudly at the podium of the first International Research Software Engineering Conference held in the Southern Hemisphere, a historic moment for both her career and the global RSE community. Having moved to South Africa, she speaks passionately about the unique challenges researchers in the Global South face: limited access to high-performance computing, unreliable funding, and the persistent digital divide. Her keynote highlights how local innovation thrives despite constraints, often out of necessity rather than choice. The audience, a diverse mix of global experts, listens intently as she calls for equitable collaborations, not just inclusion as an afterthought.
\end{story}

@@

\begin{story}{The interaction of RSEs with AI}{RSEs_with_AI}
Kim develops advanced scientific software in tandem with an AI agent that codes, debugs, and simulates experiments alongside him. The AI proactively suggests optimisations, generates reproducible workflows, and adapts to Kim’s coding style and domain-specific needs. With this partnership, Kim accelerates his research output while maintaining full control over scientific direction and ethical oversight.
\end{story}

@@

\subsection{Ethics and social considerations}
\begin{whatis}{}{ethics}
The principles of Good Scientific Practice \cite{dfg_gsp} have been a cornerstone in the conduct of research.
But their relevance hinges on constant updates in light of how research is performed.
By 2035 we want to have achieved accepted rules on the use of AI in research tasks and software development.
Open Science and Open Research are well-established in academia, and supported by respective Open Research Software
strategies of the political actors. We also hope that Data Awareness and Data Protection are still values
that are protected.
\end{whatis}

@@

% society implications
The work being undertaken by RSEs may well have direct, or indirect,
ethical implications in areas such as social engineering \cite{s2erc,siadati2024},
climate \cite{Lannelongue2023} and more.
Another example might be that they are developing code that can also be used for military purposes, such as numerical solvers or computer vision systems.
Like engineers, RSEs take the responsibility for the software they develop.
RSEs need to be aware of these issues and develop their own ``red lines''.
With the advantages of \autoref{sec:ai} in place, ever-more AI created artifacts will be in the world.
Therefore we expect a discussion within societies about their acceptable use to take place.

@@

\subsection{How have academic Institutions transformed by then?}
\begin{whatis}{}{academic}
We expect research software to be a central part of scientific practice on the same level as data and publications.
This will be supported by the first generations of RSE-Master graduates that find their way into research departments, and accelerate the pace of science.
Similar to RSEs, that are widely known and understood by university boards in 2035, we expect the concept of RSE units to be in the minds
of academic staff at any level.
We hope to see that universities use the presence of an RSE Department actively as a means to differentiate themselves from other universities.
We hope that the central RSE units have local RSE representatives, or even small RSE groups that handle tasks within a department to oversee specific projects \cite{Kempf2025-draft}.
We also hope, that specialised clusters form that pool the skills of certain specialised RSEs, with certain RSE units specializing in certain tasks like data management, HPC, or visualisation.
At universities which have them, we want to see that researchers naturally accept them as part of their available services.
Additionally researchers will seamlessly interact with infrastructures for using research software.
\end{whatis}
Academia, while claiming to be innovative \cite{wzvgbayern}, has organisationally often adopted time-scales for change that are more reminiscent to institutions thinking on the time scale of generations.
Nevertheless, the presence, even today, of initiatives like \emph{the Hidden Ref} \cite{hiddenref} hints at the potential for specialisation of research tasks. This gives rise to new job descriptions within the academic system such as specialised RSEs or other RTPs.
We have already seen some specialisation trends in recent years such as
the HPC-RSE and AI-RSE which are quite recent developments, whose appeal is stimulated by large public and private investments in HPC facilities and AI technologies.
But we could imagine new RSE specialisations emerging in the fields
of quantum computing, such as the Quantum-RSE, and in federated online platforms for research.
% European AI factories: https://digital-strategy.ec.europa.eu/en/policies/ai-factories
% European federation platform for HPC: https://eurohpc-ju.europa.eu/paving-way-eurohpc-federation-platform-2024-12-19_en
% European quantum computing: https://digital-strategy.ec.europa.eu/en/policies/quantum

On the other hand the existence of people with a certain specialised skillset opens up possibilities for pooling.
Universities will benefit from pooling these specialists into RSE units, and offer their services to researchers.
Hopefully, these RSE-units are endowed with enough trust by the university boards that translates into stable funding.
With longer-term funding and more trust to pick up longer term projects,
some RSEs can focus on code modernisation, clean-up, and modularisation,
which can pay off in faster and easier development in the future.
This is an opportunity that would not be possible if relying on short-term funded solo RSEs.

\emph{Which challenges will still persist}\\
The workshop participants expect that there will still be budgeting and administrative hurdles to creating
permanent and well-funded RSE teams.
Here it can only be strongly hoped, that, with the recognised importance of research software, research councils and funders
support the development with a reliable stream of funding.
This would also require a change of culture in research about the acceptance of research software.\\
While we hope that the barrier to researchers diminishes, the workshop participants
still expect some work to do. We hope that the compartmentalisation that occurs will help researchers re-prioritise their workload.
An upcoming challenge will be that, with the emergence of more dedicated RSE units,
networking between them will come into the foreground, because there should be a strong need to share best practices
among institutions with successful RSE Teams.
On a more organisational level, we expect the sharing of models on how to organise different RSEs with different areas of specialist expertise, to fit the different organisational constraints.
After that, organisations will have to come up with updated models for hiring people.
While a standardised RSE Master's degree makes it easier to develop the next generation of practitioners with the necessary technical skills, it will be interesting to see if
RSE are still often hired by word of mouth within an organisation(``I have this PhD here, before their PostDoc they would be a good fit for the RSE department...'' ) or if the RSE Master's degree ensures a hiring process on a more standardised level.

\begin{story}{Changes to the Universities}{changes}
Kim found work years ago at one of the first RSE units in Arcadia. After a couple of years he moved to the university of
Avalon which was interested in setting up their own department for RSEs in order to not fall behind the university of Arcadia.
Kim builds upon the model set forth by the University of Arcadia based on his valuable experience. Since the university
of Avalon has to follow different rules, not everything can be straightforwardly adapted. Life would be so much easier
if at least university regulations would be homogenised, no?
\end{story}

\emph{New challenges}
If we follow the ideas laid out in \cite{Kempf2025-draft}, then in addition to university wide RSE units,
there will be embedded RSEs in the departments. Shaping and developing these network bonds will be a task for the future.
While networking embedded RSEs and RSE units is essential for building robust, sustainable research software infrastructure, the question arises whether every group truly needs to host and manage its own infrastructure.
When each team or institution hosts its own systems, this often leads to siloed environments where resources, expertise, and data are isolated, limiting collaboration and creating inefficiencies. Such silos can result in duplicated efforts, higher costs, and challenges in maintaining data integrity and governance, as each group independently manages similar infrastructure and workflows. While local control offers flexibility, it also increases complexity and makes it harder to share innovations or best practices across the wider research community. To overcome these challenges, federated infrastructure models are needed, enabling shared access, interoperability, and coordinated development while still allowing for some local autonomy \cite{Jelinek2021}.
Federated approaches break down silos, promote knowledge exchange, and ensure that infrastructure investments benefit the broader ecosystem rather than just isolated groups.

\subsection{Impact of Alternative Career Paths for Researchers on the Structure of Research Institutes}
\begin{whatis}{}{career-paths}
We would like to see the emergence of more dedicated staff, i.e. other (digital) RTPs next to the RSE.
Also we hope that researchers can build careers at the interface of academia, industry, and policy (\eg in think tanks, NGOs,
or organisations that do knowledge transfer).
\end{whatis}
RSEs can either be embedded in research groups (belonging to the mid-level university staff),
or be members of larger RSE units.\cite{Kempf2025-draft}
The participants had the opinion that focusing on embedded RSEs would make onboarding new RSEs to existing projects easier.
This, however, comes with additional funding challenges,
as long-term funding can be more difficult to acquire in lower tiers of an of a research organisation (i.e. at departmental or individual research group level).
Third-party funding can mitigate this issue in the mid-term.
However, the participants considered central RSE groups more likely in the future,
due to the fact that there are already examples and activities from various universities in this direction
(examples include the universities of Jena, Heidelberg and Göttingen, as well as several universities in the UK where the concept of dedicated RSE groups began to emerge around 2013).
We expect funding in central groups to evolve from initially temporary positions to permanent positions,
as the groups prove themselves to be increasingly valuable and, hence, sustainable.
Embedded RSEs and central groups will need to network with each other,
and potentially request and coordinate support from external RSEs.

\begin{story}{New career paths}{New career paths}
Ken, though trained as a physicist, found his passion not in theory, but in keeping the quantum computers humming. He became the go-to expert for troubleshooting entangled qubits and stabilizing delicate cryogenic systems. His unique blend of physics knowledge and hands-on problem solving earned him a new title among his peers: the first Quantum RTP. In a world where theory meets reality, Ken quietly ensures the future of quantum computing stays on track.
\end{story}
\section{What is in this future for RSEs?}
%FG-NOTE: I think we can use this part to wrap things a bit up, and describe how this future then looks like.
%GC-NOTE: Shall we add another story box here? Ideas what we write here?
In the previous sections, we have tried to list some of the trends that we foresee.
The material presented here is based on the outputs of a workshop, but extended with the authors' ideas.
Nevertheless, there are some things that we declared out of scope here.
Among that is the ongoing digital transformation in wider society.
Digital competencies in the general population continue to increase, making the jump to software development easier than before.
But this digitalisation will also lead to changes in the way societies interact and therefore will translate into changes in how academia perceives itself.
We have also not delved into the interaction with industry much.
There are also some trends that we believe will stay with us over the coming years.
The differentiation of job profiles in academia, that has started with profiles such as the RSE will continue.
Societies will be facing new questions, that arise from technologies such as generally available AI systems.

>>>>

So where do RSEs fit into this picture? Even today, we expect to see RSEs providing technical skills and input
to research activities in a range of different areas, including:
high-quality, robust architecture and design for research software;
seamless integration with research and data infrastructures;
development of software for novel and emerging architectures;
application of specialist numerical and statistical methods within software;
optimisation of codes for use on high-performance computing infrastructure;
and green computing - efficient implementations of algorithms/methods and use of hardware

@@

\subsection{Definition and Evolution of RSEs}
\begin{whatis}{}{definition-evolution-rse}
The term RSE is established in the minds of university boards and leadership, as a valuable member in the academic framework.
New RSEs are trained through a specialised Master's programme, and their support is valued by researchers. We still teach researchers basic software skills, but the content of this lecture has drastically changed.
\end{whatis}
% Here we consider how the topics of RSEng will change in the face of the digitalisation of society.
% Also here: What will our topics be after version control doesn't need to be taught anymore?
% Digitalisation will also feature more heavily in the domain curricula.
In this discussions the workshop participants found that
the digital skills of domain scientists will increase,
and software will become easier to interact with,
reducing reliance on RSEs to train newcomers on the most basic tasks.
This means, that today's workshop content of proper versioning and data management could become obsolete
by using more intuitive tools.
Introductory courses can then focus on more higher-level concepts.
User training sessions might focus on more advanced topics
that cater to a smaller audience, or involve mentoring more frequently.

@@

\subsection{Development of the RSE community}
\begin{whatis}{}{development-rse-community}
National RSE communities are increasingly collaborating with local partners, fostering stronger networks across countries. Regular international conferences further amplify these connections, promoting the global exchange of ideas. The introduction of RSE Master's programmes, coupled with a shift in culture towards valuing Research Software, has significantly elevated its profile. This growing recognition has led to a surge in membership of national societies, reflecting the increasing importance of RSEng in the research landscape.
\end{whatis}
In recent years, we have seen the formation of RSE institutes and advocacy groups, such as
the UK Software Sustainability Institute\footnote{\url{https://www.software.ac.uk}},
the US Research Software Sustainability Institute\footnote{\url{https://urssi.us}},
the European Virtual Institute for Research Software Excellence\footnote{\url{https://everse.software}},
and the EOSC Opportunity Area Expert Group 7 on Research Software%
\footnote{\url{https://eosc.eu/opportunity-area-exp/oa7-research-software}}.
We expect these initiatives to help increase the visibility and accelerate
the professionalisation%
of the RSE role.
In addition, a multitude of RSE communities self-organised at the national
and regional level to support more viable RSE career paths in academia.
We can expect RSE positions to become more easily accessible, possibly through centrally-funded long-term positions at research institutions.
The community of RSEs will get a significant boost in their acceptance if RSE Master's programmes
become available at various universities.
With increased numbers, the visibility of the RSE community to industry will be enhanced.
\begin{story}{Development of RSE community}{RSEcommunity}
Kay stands proudly at the podium of the first International Research Software Engineering Conference held in the Southern Hemisphere, a historic moment for both her career and the global RSE community. Having moved to South Africa, she speaks passionately about the unique challenges researchers in the Global South face: limited access to high-performance computing, unreliable funding, and the persistent digital divide. Her keynote highlights how local innovation thrives despite constraints, often out of necessity rather than choice. The audience, a diverse mix of global experts, listens intently as she calls for equitable collaborations, not just inclusion as an afterthought.
\end{story}

@@

\begin{story}{The interaction of RSEs with AI}{RSEs_with_AI}
Kim develops advanced scientific software in tandem with an AI agent that codes, debugs, and simulates experiments alongside him. The AI proactively suggests optimisations, generates reproducible workflows, and adapts to Kim’s coding style and domain-specific needs. With this assistance, Kim accelerates his research output while maintaining full control over scientific direction and ethical oversight.
\end{story}

@@

\subsection{Ethics and social considerations}
\begin{whatis}{}{ethics}
The principles of Good Scientific Practice \cite{dfg_gsp} have been a cornerstone in the conduct of research.
But their relevance hinges on constant updates in light of how research is performed.
By 2035, we want to have achieved accepted rules on the use of AI in research tasks and software development.
Open Science and Open Research are by then well-established in academia, and supported by respective Open Research Software
strategies of the political actors. We also hope that Data Awareness and Data Protection are still values
that are protected.
\end{whatis}

@@

% society implications
The work being undertaken by RSEs may well have direct, or indirect,
ethical implications in areas such as social engineering \cite{s2erc,siadati2024},
climate \cite{Lannelongue2023} and more.
Another example might be that they are developing code that can also be used for military purposes, such as numerical solvers or computer vision systems.
As engineers, RSEs take the responsibility for the software they develop.
RSEs need to be aware of these issues and develop their own ``red lines''.
With the advantages of \autoref{sec:ai} in place, ever-more AI created artifacts will be in the world.
Therefore we expect a discussion within societies about their acceptable use to take place.

@@

\subsection{How have academic Institutions transformed by then?}
\begin{whatis}{}{academic}
We expect research software to be a central part of scientific practice on the same level as data and publications.
This will be supported by the first generations of RSE Master's graduates that find their way into research departments, and accelerate the pace of science.
Similar to RSEs, that are widely known and understood by university boards in 2035, we expect the concept of RSE units to be in the minds
of academic staff at any level.
We hope to see that universities use the presence of an RSE Department actively as a means to differentiate themselves from other universities.
We hope that the central RSE units have local RSE representatives, or even small RSE groups that handle tasks within a department to oversee specific projects \cite{Kempf2025-draft}.
We also hope that specialised clusters form, which pool the skills of certain specialised RSEs with certain RSE units specializing in certain tasks like data management, HPC, or visualisation.
At universities which have them, we want to see that researchers naturally accept them as part of their available services.
Additionally researchers will seamlessly interact with infrastructures for using research software.
\end{whatis}
Academia, while claiming to be innovative \cite{wzvgbayern}, has organisationally often adopted time scales for change that are more reminiscent to institutions thinking on the time scale of generations.
Nevertheless, the presence, even today, of initiatives like \emph{the Hidden Ref} \cite{hiddenref} hints at the potential for specialisation of research tasks. This gives rise to new job descriptions within the academic system such as specialised RSEs or other RTPs.
We have already seen some specialisation trends in recent years such as
the HPC-RSE and AI-RSE which are quite recent developments, whose appeal is stimulated by large public and private investments in HPC facilities and AI technologies.
But we could imagine new RSE specialisations emerging in the fields
of quantum computing, such as the Quantum-RSE, and in federated online platforms for research.
% European AI factories: https://digital-strategy.ec.europa.eu/en/policies/ai-factories
% European federation platform for HPC: https://eurohpc-ju.europa.eu/paving-way-eurohpc-federation-platform-2024-12-19_en
% European quantum computing: https://digital-strategy.ec.europa.eu/en/policies/quantum

On the other hand, the existence of people with a certain specialised skill set opens up possibilities for pooling.
Universities will benefit from pooling these specialists into RSE units, and offer their services to researchers.
Hopefully, these RSE-units are endowed with enough trust by the university boards that translates into stable funding.
With longer-term funding and more trust to pick up longer term projects,
some RSEs can focus on code modernisation, clean-up, and modularisation,
which can pay off in faster and easier development in the future.
This is an opportunity that would not be possible if relying on short-term funded solo RSEs.

\emph{Which challenges will still persist}\\
The workshop participants expect that there will still be budgeting and administrative hurdles to creating
permanent and well-funded RSE teams.
Here it can only be strongly hoped that, with the recognised importance of research software, research councils and funders
support the development with a reliable stream of funding.
This would also require a change of culture in research about the acceptance of research software.\\
While we hope that the barrier to researchers diminishes, the workshop participants
still expect some work to do. We hope that the compartmentalisation that occurs will help researchers re-prioritise their workload.
An upcoming challenge will be that, with the emergence of more dedicated RSE units,
networking between them will come into the foreground, because there should be a strong need to share best practices
among institutions with successful RSE Teams.
On a more organisational level, we expect the sharing of models on how to organise different RSEs with different areas of specialist expertise, to fit the different organisational constraints.
After that, organisations will have to come up with updated models for hiring people.
While a standardised RSE Master's degree makes it easier to develop the next generation of practitioners with the necessary technical skills, it will be interesting to see if
RSE are still often hired by word of mouth within an organisation(``I have this PhD here, before their PostDoc they would be a good fit for the RSE department...'' ) or if the RSE Master's degree ensures a hiring process on a more standardised level.

\begin{story}{Changes to the Universities}{changes}
Kim found a job years ago at one of the first RSE units in Arcadia. After a couple of years, Kim moved to the university of
Avalon which was interested in setting up their own department for RSEs in order to not fall behind the university of Arcadia.
Kim builds upon the model set forth by the University of Arcadia based on his valuable experience. Since the university
of Avalon has to follow different rules, not everything can be straightforwardly adapted. Life would be so much easier
if at least university regulations would be homogenised, no?
\end{story}

\emph{New challenges}
If we follow the ideas laid out in related community work~\cite{Kempf2025-draft}, then in addition to university wide RSE units,
there will be embedded RSEs in the departments. Shaping and developing these network bonds will be a task for the future.
While networking embedded RSEs and RSE units is essential for building robust, sustainable research software infrastructure, the question arises whether every group truly needs to host and manage its own infrastructure.
When each team or institution hosts its own systems, this often leads to siloed environments where resources, expertise, and data are isolated, limiting collaboration and creating inefficiencies. Such silos can result in duplicated efforts, higher costs, and challenges in maintaining data integrity and governance, as each group independently manages similar infrastructure and workflows. While local control offers flexibility, it also increases complexity and makes it harder to share innovations or best practices across the wider research community. To overcome these challenges, federated infrastructure models are needed, enabling shared access, interoperability, and coordinated development while still allowing for some local autonomy \cite{Jelinek2021}.
Federated approaches break down silos, promote knowledge exchange, and ensure that infrastructure investments benefit the broader ecosystem rather than just isolated groups.

\subsection{Impact of Alternative Career Paths for Researchers on the Structure of Research Institutes}
\begin{whatis}{}{career-paths}
We would like to see the emergence of more dedicated staff, i.e. other (digital) RTPs next to the RSE.
Also we hope that researchers can build careers at the interface of academia, industry, and policy (\eg in think tanks, NGOs,
or organisations that do knowledge transfer).
\end{whatis}
RSEs can either be embedded in research groups (belonging to the mid-level university staff),
or be members of larger RSE units.\cite{Kempf2025-draft}
The participants had the opinion that focusing on embedded RSEs would make onboarding new RSEs to existing projects easier.
This, however, comes with additional funding challenges,
as long-term funding can be more difficult to acquire in lower tiers of an of a research organisation (i.e., at departmental or individual research group level).
Third-party funding can mitigate this issue in the mid-term.
However, the participants considered central RSE groups more likely in the future,
due to the fact that there are already examples and activities from various universities in this direction
(examples include the universities of Jena, Heidelberg, and Göttingen, as well as several universities in the UK where the concept of dedicated RSE groups began to emerge around 2013).
We expect funding in central groups to evolve from initially temporary positions to permanent positions,
as the groups prove themselves to be increasingly valuable and, hence, sustainable.
Embedded RSEs and central groups will need to network with each other,
and potentially request and coordinate support from external RSEs.

\begin{story}{New career paths}{New career paths}
Ken, though trained as a physicist, found his passion not in theory, but in keeping the quantum computers humming. He became the go-to expert for troubleshooting entangled qubits and stabilizing delicate cryogenic systems. His unique blend of physics knowledge and hands-on problem solving earned him a new title among his peers: the first Quantum RTP. In a world where theory meets reality, Ken quietly ensures that the future of quantum computing stays on track.
\end{story}
\section{What is in this future for RSEs?}
%FG-NOTE: I think we can use this part to wrap things a bit up, and describe how this future then looks like.
%GC-NOTE: Shall we add another story box here? Ideas what we write here?
In the previous sections, we have tried to list some of the trends that we foresee.
The material presented here is based on the outputs of a workshop, but extended with the authors' ideas.
Nevertheless, there are some things that we declared out of scope here.
Among that is the ongoing digital transformation in wider society.
Digital competencies in the general population continue to increase, making the jump to software development easier than before.
But this digitalisation will also lead to changes in the way societies interact and therefore will translate into changes in how academia perceives itself.
We have also not sufficiently delved into the interaction with industry.
There are also some trends that we believe will stay with us over the coming years.
The differentiation of job profiles in academia, that has started with profiles such as the RSE will continue.
Societies will be facing new questions, arising from technologies such as generally available AI systems.
